---
title: "Assignment 6: Using Sobol with an ODE"
author: "Nico Gavigan"
format: 
  pdf:
    toc: true
execute: 
  warning: false
  message: false
df-print: kable
---

# Using Sobol with an ODE 

```{r, echo=FALSE}
library(sensitivity)
library(tidyverse)
library(deSolve)
library(lhs)
library(purrr)
library(here)
library(ggpubr)
```

#### Run Forest Growth Model

```{r}
# Source forest growth model
source(here("R/dcarbon.r"))
dcarbon

# Create parameters and set conditions
parms = list(thresh = 50, K = 250, r = 0.01, g = 2)
initialC = 10
time = seq(from = 1, to = 300, by = 5)

# run forest growth model
forestgrowth = ode(func = dcarbon, y = initialC, times = time, parms = parms)
```

#### Plot model output

```{r}
# Assign column names
colnames(forestgrowth) <- c("time", "carbon")

# Plot model output
ggplot(as.data.frame(forestgrowth), aes(time, carbon)) +
  geom_line() +
  labs(y = "Forest Size (kg C)",
       x = "Time") +
  theme_minimal() 
```

## Begin Sobol Global Sensitivity Analysis

```{r}
# Sobol Global Sensitivity Analysis 
# Create X1 data frame
np = 300
thresh = rnorm(mean = 50, sd = 50*0.1, n = np)
K = rnorm(mean = 250, sd = 250*0.1, n = np)
r = rnorm(mean = 0.01, sd = 0.01*0.1, n = np)
g = rnorm(mean = 2, sd = 2*0.1, n = np)

X1 = cbind.data.frame(thresh = thresh, K = K, r = r, g = g)

# Create X2 data frame
np = 300
thresh = rnorm(mean = 50, sd = 50*0.1, n = np)
K = rnorm(mean = 250, sd = 250*0.1, n = np)
r = rnorm(mean = 0.01, sd = 0.01*0.1, n = np)
g = rnorm(mean = 2, sd = 2*0.1, n = np)

X2 = cbind.data.frame(thresh = thresh, K = K, r = r, g = g)

# Run SobolSalt
sensCarbon = sobolSalt(model = NULL, X1, X2, nboot = 300)
colnames(sensCarbon$X) = c("thresh", "K", "r", "g")

head(sensCarbon$X)
```

```{r, echo=FALSE}
# Checking to make sure it works 
initialC <- 10
parms <- list(thresh = sensCarbon$X[1, 1], K = sensCarbon$X[1, 2], r = sensCarbon$X[1, 3], 
              g = sensCarbon$X[1,4])

res <- ode(y = initialC, times = time, func = dcarbon, parms = parms)
res <- as.data.frame(res)
colnames(res) <- c("time", "C")

ggplot(res, aes(time, C)) +
  geom_line()
```

## How does maximum forest size over 300 years vary with these parameters?

#### Create Metric Function

```{r}
max_C = function(result){
  max(result$C)
}
```

#### Create wrapper function

```{r}
# Wrapper function
p_wrapper <- function(thresh, K, r, g, time, initialC, odefunc, parms, metricfunc) {
  parms <- list(thresh = thresh, K = K, r = r, g = g)
  result <- ode(y = initialC, times = time, func = odefunc, parms = parms)
  result <- as.data.frame(result)
  colnames(result) <- c("time", "C")
  # get metrics
  metrics <- metricfunc(result)
  return(metrics)
}
```

#### Extract Results 

```{r}
allresults <- as.data.frame(sensCarbon$X) %>% pmap_dbl(p_wrapper, initialC = initialC, 
                                                       time = time, odefunc = dcarbon,
                                                       metricfunc = max_C)
allresults_df <- data.frame(max_carbon = allresults)

head(allresults_df)
```

#### Plot Results

```{r}
ggplot(allresults_df, aes(y = max_carbon)) +
  geom_boxplot() +
  labs(y = "Maximum Forest Size") +
  theme_minimal()
```

#### Create Sobol Indicies

```{r}
sensCarbon_300 = tell(sensCarbon, allresults)

rownames(sensCarbon_300$S) = c("thresh", "K", "r", "g")
sensCarbon_300$S
rownames(sensCarbon_300$T) = c("thresh", "K", "r", "g")
sensCarbon_300$T
```

------------------------------------------------------------------------

The Sobol first and total order indicies reveal that r, the early exponential growth rate and K, the carrying capacity are the top two most important variables to forest growth. In the context of climate change, factors that impact growth rate (eg. temperature, fertilization, water availability) will have the most significant impact on forest growth. Factors affecting carrying capacity, such as wildfire disturbances and max carbon storage, will have the second greatest impact on forest growth.
