---
title: "ESM 232 - Lab 1"
author: "Nico Gavigan"
format: html
editor: visual
---

# Notes

Common practice to have function name the same as the data file name

There is a style guide in R that shows tidyverse best practices for functions.

#### Example of function in Assignment 1

```{r}
# solar_energy.R          # function "arguments"
solar_energy <- function(A, r = 0.2, H, PR = 0.75) {
  
  # Calculate and return E
  E = A * r * H * PR
  # Returns can be explicit but it returns E anyways in this case 
  return(E) # I think this goes here but might be outside of brackets
}

# Ex of what not to do
solar_energy <- function(A, r, H, PR) {
  r = 0.2    # this is called "hard-coding" a value and is not recommended
  PR = 0.75
  
  E = A * r * H * PR
}
```

## Sensitivity Analysis with ethresh

-   Assume +- 15% uncertainty in ethresh parameter (uniform distribution)

-   Default values (from the function) is 10,000

-   Code so that it is easy to change the uncertainty

-   Think about a data structure to allow us to see how uncertainty impacts annual power estimation for each year, or mean annual power generation

```{r}
# Source the model
source(here("R/solarpv.R"))

# read in R formatted data
load(here("Data/sierraczosolar_clean.rda"))
```

```{r}
library(tidyverse)
library(here)
library(purrr)
```

```{r}
# Assume +-15% uncertainty in ethresh parameter (uniform distribution)

ethresh <- rnorm(mean = 10000, sd = 0.15*10000, n = 20)

site3 <- ethresh |> map(~ solarpv(area = 0.1, solar = sierraczosolar, etype = "direct", 
                                  ethresh = .x))

plot(ethresh)

head(site3)
print(site3)
```

```{r}
# Ojas method
deviation <- 0.15
base_thresh <- 10000
ethresh <- runif(
  min = base_thresh - deviation * base_thresh,
  max = base_thresh + deviation * base_thresh, n = 40
)

plot(ethresh)
```

```{r}
 tmp <0 map_df(site2, '[', c("annual"))
 site2df <- data.frame(
   year = tmp$annual$year
   elec = tmp$annual$year
 )
```
